---
title: 'Reproducible Research: Peer Assessment 1'
author: "Dhruvin Shah"
date: "5/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
1. Loading data using read.csv() function

```{r}
data_raw <- read.csv("activity.csv", header = TRUE)
head(data_raw)
```

Dimension of Data:
```{r}
dim(data_raw)
```

Structure of Data:
```{r}
str(data_raw)
```

2. Preprocessing the data

```{r}
data <- data_raw
data$date <- as.Date(data$date)
str(data)
```


## What is mean total number of steps taken per day?

Finding total number of steps by using group_by() and summarise() function of dplyr library.
```{r}
library(dplyr)
TotalStepsData <- data %>%
    na.omit() %>%
    group_by(date) %>%
    summarise(Steps = sum(steps))
TotalStepsData <- as.data.frame(TotalStepsData)
head(TotalStepsData)
```

Histogram of Total Steps is
```{r}
hist(TotalStepsData$Steps, 
     main = "Histogram of Total Number of Steps",
     breaks = 25,
     col = "#086A87",
     xlab = "Step")
```

Mean of Total Steps is:
```{r}
mean(TotalStepsData$Steps)
```

Median of Total Steps is:
```{r}
median(TotalStepsData$Steps)
```

## What is the average daily activity pattern?

Finding Average steps during the 5 - min interval using group_by() and summarise() functions of dplyr library.
```{r}
AverageStepsData <- data %>%
    na.omit() %>%
    group_by(interval) %>%
    summarise(Steps = mean(steps))
AverageStepsData <- as.data.frame(AverageStepsData)
head(AverageStepsData)
```

Plotting the Time-series using ggplot() function
```{r}
library(ggplot2)
ggplot(data = AverageStepsData, 
       aes(x = interval,
           y = Steps)) +
    geom_line(color = "#00AFBB", size = 0.5) +
    labs(title = "Average Number of Steps Taken in an Interval") + 
    xlab("5 minute Interval") +
    ylab("Average Steps Taken")
```


## Imputing missing values

Checking for Missing Values:
```{r}
any(is.na(data))
```

Total Number of Missing Values:
```{r}
sum(is.na(data))
```

Missing Values in each column:
```{r}
sapply(data, function(x) sum(length(which(is.na(x)))))
```

Percentage of Missing Values:
```{r}
sapply(data, function(x) (mean(is.na(x)))*100)
```

Replacing missing values with average of interval
```{r}
cleandata <- data
 for(i in AverageStepsData$interval) {
     cleandata[cleandata$interval == i & is.na(cleandata$steps), ]$steps <- AverageStepsData$Steps[AverageStepsData$interval == i]
 }
```

Checking for any NA values in updated dataset
```{r}
any(is.na(cleandata))
head(cleandata)
```

Finding total number of steps in clean data by using group_by() and summarise() function of dplyr library.
```{r}
TotalStepsCleanData <- cleandata %>%
    group_by(date) %>%
    summarise(steps = sum(steps))
TotalStepsCleanData <- as.data.frame(TotalStepsCleanData)
head(TotalStepsCleanData)
```

Histogram of Total Steps of Clean Data is
```{r}
hist(TotalStepsCleanData$steps, 
     main = "Histogram of Total Number of Steps",
     breaks = 25,
     col = "#086A87",
     xlab = "Step")
```

Mean of Total Steps is:
```{r}
mean(TotalStepsCleanData$steps)
```

Median of Total Steps is:
```{r}
median(TotalStepsCleanData$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

Determing the Phase of Week using weekdays() function
```{r}
cleandata$phase <- ""
cleandata$date <- as.Date(cleandata$date)
cleandata[weekdays(cleandata$date) == "Saturday" | weekdays(cleandata$date) == "Saturday",]$phase <- "weekend"
cleandata[!(weekdays(cleandata$date) == "Saturday" | weekdays(cleandata$date) == "Saturday"),]$phase <- "weekday"
head(cleandata)
```

Finding Average steps of Week Phase during the 5 - min interval using group_by() and summarise() functions of dplyr library.
```{r}
AverageStepsWeekPhase <- cleandata %>%
    group_by(phase, interval) %>%
    summarise(steps = mean(steps))
AverageStepsWeekPhase <- as.data.frame(AverageStepsWeekPhase)
head(AverageStepsWeekPhase)
```

Subsetting the Weekday and Weekend Data from Weekphase
```{r}
AverageStepsWeekday <- AverageStepsWeekPhase %>%
    filter(phase == "weekday")
head(AverageStepsWeekday)
```

```{r}
AverageStepsWeekend <- AverageStepsWeekPhase %>%
    filter(phase == "weekend")
head(AverageStepsWeekend)
```

Plotting the Time Series of Weekdays and Weekends Activity
```{r}
par(mfrow = c(1,2))
plot(AverageStepsWeekday$interval, AverageStepsWeekday$steps,
     main = "Weekday",
     xlab = "5 minute Interval",
     ylab = "Average Steps Taken",
     type = "l", lwd = 2, col = "navy")

plot(AverageStepsWeekend$interval, AverageStepsWeekend$steps,
     main = "Weekend",
     xlab = "5 minute Interval",
     ylab = "Average Steps Taken",
     type = "l", lwd = 2, col = "red")
```

```{r}

```
